---
title: "GCA Analysis Notebook"
author: "Bud Talbot, Derek Briggs"
output:
  html_notebook: default
  html_document: default
---

##Load data
Load data from GCA_ALL_DATA_fa15_sp16_fa16_03302017, name the new data frame FSF_alldata (FSF = "Fall Spring Fall")
```{r}
library(readr)
FSF_alldata <- read_csv("C:/Users/talbotr/Desktop/Keck/FSF_alldata.csv")
```

##Subset data
This code subsets GCA pre and post responses from FSF all data- including cl, Sex and ethnicity codes
```{r}
GCA_prepost <- subset(FSF_alldata, select=c(StudentID, Semester, cl, Pre1:Post25, Sex:Demo4White5))
```


Use lapply to find means of all GCA pre and post item responses and write to new list. Also need to do some basic item-total correlations here. This code is not implemented yet
```{r}
GCA_item_Means <- lapply(3:52, mean)
```

##Calculate GCA Pre and Post total scores
This code calculates GCA pre and post total scores for each case adds the new vars to  FSF_alldata
```{r}
Pre_tot <- rowSums(GCA_prepost[4:28], na.rm = TRUE)
Post_tot <- rowSums(GCA_prepost[29:53], na.rm = TRUE)
GCA_prepost$Pre_tot <- Pre_tot
GCA_prepost$Post_tot <- Post_tot
```

##Calculate raw gain and <g>
This code calculates raw gain and <g> for each case and appends the new vars to FSF_alldata
```{r}
raw_gain <- (GCA_prepost[62] - GCA_prepost[61])
norm_gain <- raw_gain/(25 - GCA_prepost[61]) 
GCA_prepost$gain <- raw_gain
GCA_prepost$norm_gain <- norm_gain
```
Could follow this up with some scatterplots or histograms of gains within each class


##Filter and calculate class <g> and ES
Now filter (using dplyr) FSF_alldata for cl = 1 and Semester = Fa2015 from GCA_prepost, then calculate ES and g (based on averaging GCA pre and psot scores- not based on averaging individual <g> and ES) for this class (class 1 from Fall 2015). 
```{r}
library(dplyr)
cl1f15 <- filter(GCA_prepost, Semester=="Fa2015", cl==1)
cl1f15_ES <- (mean(cl1f15$Post_tot, na.rm = TRUE) - mean(cl1f15$Pre_tot, na.rm = TRUE))/sd(cl1f15$Pre_tot, na.rm = TRUE)
cl1f15_g <- (mean(cl1f15$Post_tot, na.rm = TRUE) - mean(cl1f15$Pre_tot, na.rm = TRUE))/(25-mean(cl1f15$Pre_tot, na.rm = TRUE))
cl1f15gains <- c(cl1f15_ES, cl1f15_g)
print(cl1f15gains)
```

NOTE ON ABOVE: CHANGE <g> TO NORM CHANGE c ONCE THE FUNCTION IS WORKING PROPERLY

I need to make the above code a function and then use lapply(?) to iterate over all classes and Semesters, or just increment and loop it over and over. Then write each resulting vector (e.g. cl1f15gains) to a new table. Thoughts on how to code that nicely?


##Plot histogram of pre and post GCA for a class
Plot a histogram of class 1 Fall 2015 pre and post GCA scores
```{r}
library(ggplot2)
library(RColorBrewer)
df <-  data.frame(x = cl1f15$Pre_tot, x2 = cl1f15$Post_tot)

g <-  ggplot(df, aes(x)) + geom_histogram(aes(x = x, y = ..count..),
                  binwidth = 1, color="black", fill="blue", alpha=0.5) + 
  geom_histogram( aes(x = x2, y = -..count..), 
                  binwidth = 1, color="black", fill= "green", alpha=0.5) +
  scale_x_continuous(name = "GCA Pre/Post Score", 
                     limits=c(0, 25)) +
  scale_y_continuous(name = "Count") +
  ggtitle("Class 1 Fall 2015 GCA Pre and Post Scores")
print(g)
```

Or this if we don't want the post distribution inverted
```{r}
g2 <-  ggplot(df, aes(x)) + geom_histogram(aes(x = x, y = ..count..),
                  binwidth = 1, color="black", fill="blue", alpha=0.5) + 
  geom_histogram( aes(x = x2, y = ..count..), 
                  binwidth = 1, color="black", fill= "green", alpha=0.5) +
  scale_x_continuous(name = "GCA Pre/Post Score", 
                     limits=c(0, 25)) +
  scale_y_continuous(name = "Count") +
  ggtitle("Class 1 Fall 2015 GCA Pre and Post Scores")
print(g2)
```


##Local item <g> and ES
Next we need to calculate pre and post totals and gains (<g>, ES) for the local items for each class (note that this will *not* be for Fa15 or Sp16, only Fa16 and Sp17). 

##GCA and local gain scores by learning goals
Then we need to code new vars for GCA and local items grouped by learning goal, and find gains for these item groupings.

